The importance of measurement uncertainty in ecological management

### 2017-12-22 appendix/appendix.Rmd
Appendix example based on f(x - h); recruit before harvest.

write_csv(sims, "sims.csv")
```

@@

@@

policy_table %>% 
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))

Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.  The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles.

@@

## Historical data estimation

## Historical data POMDP & other solutions

## Historical combined plot

>>>>

write_csv(sims, "appendix_files/sims.csv")
```

@@

write_csv(policy_table, "appendix_files/policy_table.csv")
```

@@

```{r}
policy_table %>% 
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line()  + 
  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
```

Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.  The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles when we plot in terms of escapement instead of harvest (as is common in the optimal control literature in fisheries, e.g. @Sethi2005).

Note that under the optimal solution, escapement is effectively constant at $B_{MSY} = 0.5$: for all states above a certain size the population
is harvested back down to that size. Note that even stocks observed at states slightly below $B_{MSY} = 0.5$ achieve this target escapement, since we are following the classic Graham Shaeffer formulation here where we observe first, then recruitment happens before harvest, and thus we see the population at smaller size than we harvest it.  In classical escapement analysis, observations are usually indexed instead to occur immediately before harvests, and this inflection point occurs right at $B_{MSY}$.

@@

In the above plots for the MSY

```{r}
unif_prior = rep(1, length(states)) / length(states) # initial belief
unif <- compute_policy(alpha, m$transition, m$observation, m$reward,  unif_prior)

```{r}
i_star <- which.min(abs(states - S_star))
 i_low <- which.min(abs(states - 0.5 * S_star))
 i_high <- which.min(abs(states - 1.5 * S_star))
#i_low <- 2
#i_high <- which.min(abs(states - K))

prior_star <- m$observation[,i_star,1]
prior_low <- m$observation[,i_low,1]
prior_high <- m$observation[,i_high,1]

star <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_star)
low <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_low)
high <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_high)

### 2017-12-22 appendix/appendixB.Rmd
Appendix example based on f(x - h); recruit before harvest.

```{r libraries}
# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
library(tidyverse)
library(sarsop)
library(nimble)
library(parallel)
```

```{r}
download.file("https://depts.washington.edu/ramlegac/wordpress/databaseVersions/RLSADB_v3.0_(assessment_data_only)_excel.zip",
              "ramlegacy.zip")
```

@@

```{r}
sims <- pomdp_sims %>% 
  mutate(optimal = states[optimal] * N) %>% # original scale
  select(year, optimal, commonname, sigma_m) %>%
  rename(stock = optimal) %>% 
  ## treat each sigma_m value as separate 'model'
  mutate(sigma_m = as.factor(sigma_m)) %>%
  mutate(model = recode(sigma_m, 
                        "0" = "reed",
                        "0.1" = "pomdp_0.1", 
                        "0.2" = "pomdp_0.2")) %>%
  select(-sigma_m) %>%
  bind_rows(historical)
```

```{r}
sims %>%
  filter(model %in% c("biomass", "catch", "pomdp_0.1", "reed", "pgy")) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  scale_color_manual(values = c("black", "grey", "#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")) +
  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
```

>>>>

```{r libraries, message=FALSE}
# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
library(tidyverse)
library(sarsop)
library(nimble)
library(parallel)
```

```{r}
if(!file.exists("ramlegacy.zip")){
download.file("https://depts.washington.edu/ramlegac/wordpress/databaseVersions/RLSADB_v3.0_(assessment_data_only)_excel.zip",
              "ramlegacy.zip")
}
```

@@

```{r}
sims <- pomdp_sims %>% 
  mutate(optimal = states[optimal] * N) %>% # original scale
  select(year, optimal, commonname, sigma_m) %>%
  rename(stock = optimal) %>% 
  ## treat each sigma_m value as separate 'model'
  mutate(sigma_m = as.factor(sigma_m)) %>%
  mutate(model = recode(sigma_m, 
                        "0" = "Reed",
                        "0.1" = "POMDP", 
                        "0.2" = "POMDP_0.2")) %>%
  select(-sigma_m) %>%
  bind_rows(historical)

write_csv(sims,"appendixB_files/sims.csv")
```

```{r}
sims %>%
  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  scale_color_manual(values = c("black", "grey", "#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")) +
  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
```

### 2017-12-22 appendix/appendix_harvestfirst.Rmd
Appendix example based on f(x - h); recruit before harvest.

```{r helper_routines, include = FALSE}
## map cts values to nearest grid indices 
index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))

det_simulations <- function(x0 = 1/6, Tmax = 20, 
                              B_MSY = K / 2, 
                              F_MSY = r / 2 , 
                              F_PGY = 0.8 * F_MSY){

  msy_policy <- function(x) F_MSY * x
  pgy_policy <- function(x) F_PGY * x
  det_policy <- function(x) pmax(f(x,0) - B_MSY,0) ## ASSUMES harvest takes place *after* recruitment
  reed_policy  <- function(x) pmax(x - B_MSY,0) ## Optimal when harvest takes place *before* recruitment, assumes we are seeing a pre-harvest size.
  ## consequently, if we use such a policy when we are seeing a the post-harvest size, we give too much buffer.

  # a simulation function
  det_sim <- function(policy, f, x0, Tmax){
    action <- state <- obs <- as.numeric(rep(NA,Tmax))
    state[1] <- x0
    for(t in 1:(Tmax-1)){
      obs[t] <- state[t]

      action[t] <- policy(obs[t])
      state[t+1] <- f(state[t], action[t]) 
    }
    data.frame(time = 1:Tmax, state, action, obs)
  }

  # simulate each policy, returning all results in a single unified data frame 
  sims <- 
    list(msy = msy_policy, 
         det = det_policy, 
         pgy = pgy_policy
         #opt = reed_policy
         ) %>% 
    map_df(det_sim, 
           f, x0, Tmax, 
           .id = "method")

  sims
}
```

@@

## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
f <- function(x, h){ 
    x + x * r * (1 - x / K) - pmin(x,h)
}
```

@@

If we observe->recruit->harvest, $X_{t+1} = f(X_t) - h_t$, then we see the stock right after a harvest and before recuitment, when it is at $X_t ~ $B_{MSY}$.  
If we instead did observe->harvest->recruit, we would see the stock at its pre-harvest size, $X_t ~ B_{MSY} + $H_{MSY}$.

```{r}
B_MSY <- S_star
MSY <- f(B_MSY,0) - B_MSY

@@

```{r optional_checks}
## Make sure these match the known analytic solutions for Graham-Schaefer model (only true of discount = 1)
if(discount == 1){
  testthat::expect_equal(B_MSY, K / 2)
  testthat::expect_equal(MSY, r * K / 4)
  testthat::expect_equal(F_MSY, r / 2)
}
```

@@

```{r}
det_simulations(x0 = 1/6, Tmax = 20, B_MSY, F_MSY, F_PGY) %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1) 
```

@@

Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.  The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles when we plot in terms of escapement instead of harvest (as is common in the optimal control literature in fisheries, e.g. @Sethi2005).

Note that under the optimal solution, escapement is effectively constant at $B_{MSY} = 0.5$: for all states above a certain size the population
is harvested back down to that size. Note that even stocks observed at states slightly below $B_{MSY} = 0.5$ achieve this target escapement, since we are following the classic Graham Shaeffer formulation here where we observe first, then recruitment happens before harvest, and thus we see the population at smaller size than we harvest it.  In classical escapement analysis, observations are usually indexed instead to occur immediately before harvests, and this inflection point occurs right at $B_{MSY}$.

@@

In the above plots for the MSY

```{r}
unif_prior = rep(1, length(states)) / length(states) # initial belief
unif <- compute_policy(alpha, m$transition, m$observation, m$reward,  unif_prior)

```{r}
i_star <- which.min(abs(states - S_star))
 i_low <- which.min(abs(states - 0.5 * S_star))
 i_high <- which.min(abs(states - 1.5 * S_star))
#i_low <- 2
#i_high <- which.min(abs(states - K))

prior_star <- m$observation[,i_star,1]
prior_low <- m$observation[,i_low,1]
prior_high <- m$observation[,i_high,1]

star <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_star)
low <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_low)
high <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_high)

>>>>

@@

## Unlike classic Graham-Schaefer, this assumes harvest occurs right after assessment, before recruitment
f <- function(x, h){ 
    s <- pmax(x-h, 0) 
    s + s * r * (1 - s / K) 
}
```

@@

Since we `observe` -> `harvest` -> `recruit`, we would observe the stock at its pre-harvest size, $X_t \sim B_{MSY} + H_{MSY}$.

```{r}
#B_MSY <- S_star     # recruit first, as in classic Graham-Schaefer
B_MSY <- f(S_star,0) # harvest first

#MSY <- f(B_MSY,0) - B_MSY  # recruit first
MSY <- B_MSY - S_star  # harvest first

@@

@@

msy_policy <- function(x) F_MSY * x
pgy_policy <- function(x) F_PGY * x

## ASSUMES harvest takes place *before* recruitment, f(x-h), not after.
escapement_policy <- function(x) pmax(x - S_star,0)  
x0 <- K/6
Tmax = 20
det_sim <- function(policy, f, x0, Tmax){
    action <- state <- obs <- as.numeric(rep(NA,Tmax))
    state[1] <- x0

    for(t in 1:(Tmax-1)){
      action[t] <- policy(state[t])
      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
      state[t+1] <- f(state[t], action[t]) 
    }
    data.frame(time = 1:Tmax, state, action, obs)
  }

sims <- 
  list(msy = msy_policy, 
       pgy = pgy_policy,
       det = escapement_policy) %>% 
  map_df(det_sim, 
         f, x0, Tmax, 
         .id = "method") 
sims %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1)

@@

Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.  The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles.

@@

### 2017-12-22 manuscripts/manuscript.Rmd
Appendix example based on f(x - h); recruit before harvest.

bibliography: citations.bib
output: rticles::elsevier_article
---

@@

- Replicate simulations

@@

- historical data.

```{r}
```

@@

>>>>

bibliography: citations.bib
output: rticles::elsevier_article
layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column

---

```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
library(tidyverse)
library(gridExtra)
library(ggthemes)
colours <- ptol_pal()(2)
```

```{r suggested_libs}
library(hrbrthemes)
library(extrafont)
##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
hrbrthemes::import_roboto_condensed()     # One-time use to set up fonts 
extrafont::loadfonts()
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
#ggplot2::theme_set(theme_bw()) # simpler theme if you do not have these fonts
```

@@

### Replicate simulations

@@

### Comparisons to historical data

```{r}
sims <- read_csv("../appendix/appendixB_files/sims.csv")
sims %>%
  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  scale_color_manual(values = c("black", "grey", "#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")) +
  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
```

@@

# References
